Demo arXiv template

### 2023-07-04 tidyindex.qmd
outline for introducing done

<!-- Each individual index follows its own data pipeline and it can be difficult to evaluate how an index can be affected by tweaking parameters in a certain step, rearranging the order of steps, using to different method.  -->

<!-- However, even if most of the indexes have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember, classify, and compare between. -->

@@

Many concepts relevant to decision making cannot be directly measured, however, they are crucial for resource allocation, early prevention, and other operational purpose. 
For example, fire authorities would be interested to quantify fire risk since bushfires can have a huge impact on monetary loss, health, and the local ecosystem.
Climatologists would be interested in monitoring the change in global climate since variability in atmospheric and oceanic conditions has a direct impact on global weather and climate.
Usually this concept of interest is associated with more than one variables and these variables need to be integrated to make decisions on the subject matter.
<!-- From a modelling perspective, these unobserved quantities are called latent variables and existing methods can incorporate these latent variables in the model structure.  -->
A common approach to quantify concepts like these is to construct an index using these relevant variables. 
This allows researchers to compare the quantity of interest across entities (i.e. countries, regions) and also cross time.

@@

In this article, an index is defined as a tool to quantify a concept of interest that does not have a direct measure.
The concept of interest doesn't have a direct measure can because it is impractical to measure at the population level. 
For example, it would be nearly impossible to include all the available stocks in the market to characterise stock market behavior, so indexes like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite select a representative set of stocks to measure the overall market behavior. 
Also belonging to this category are the economic indexes like the Consumer Price Index, where price changes of a basket of items are weighted to measure inflation. 
The lack of direct measure could also because the concept itself is an unobservable human construction, rather than a physical quantity that can be measured.
Many natural hazard and social concepts falls into this category. 
This includes drought indexes constructed from meteorological, agricultural, hydrological, and social-economic variables, e.g. Standardised Precipitation Index (SPI) [@mckee1993relationship] and Aggregated Drought Index (ADI) [@keyantash_aggregate_2004] among others.
Social development indexes like Human Development Index [@hdi] and Global Liveability Index [@gli] measure various aspects of the quality of human capital and urban life.

*still need to tweak the tone a bit: "they are called index, they are not the index we will talk about"*

Despite many quantity having the term *index* in their name, they cannot be technically classified as indexes according to the definition given above. 
The reason for these quantities to lose their index memberships is that they are variables can be accurately measured given the instrument precision. 
This includes quantities like precipitation of the driest month or percentage of days when maximum temperature is below 10th percentile. 
They are measures of precipitation and percentage of days under specific conditions (dries month, maximum temperature below 10th percentile).
They are variables, or indicators, that can be used to construct indexes but are not indexes themselves. 
Similarly, a set of remote sensing indexes are not indexes, since they are measures of electromagnetic wave reflectance.
This includes Normalized Difference Vegetation Index (NDVI) [@tucker_red_1979], derived from the ratio of difference over sum on two segments in the spectrum, also called band: near-infrared (NIR) and red.
So are the "indexes" derived from NDVI, e.g. Vegetation Condition Index [@kogan_application_1995]. 
Notice that this does not exclude all the construction derived from remotes sensor variables to be valid indexes.
For example, Vegetation Drought Response Index [@brown_vegetation_2008] is a valid index since it integrates climate, satellite, and biophysical variables to quantify vegetation stress.

>>>>

@@

Many concepts relevant to decision making cannot be directly measured, however, they are crucial for resource allocation, early prevention, and other operational purpose.

  - USDM and European Drought Observatory publishes drought indicator to [...], 
  - NOAA and BOM producing forecast on El Nino alert system to [...]
  - OECD and UN producing indexes measuring socio-economic aspects of country for policy making.

There decision making often needs to consider multiple aspects and indexes is a statistical tool to pack these information into a single number, making it easier for policy makers and public communication.

@@

  - Indexes can be composed from single variables or multiple variables (composite index)
  - to make the definition even more complex is that indexes themselves are used with other variables to compose indexes (many drought indexes using an index called SPI). 
  - Some indexes may be more simple than you think:  
    - a single transformation (NDVI [@tucker_red_1979] takes two spectrum: red and near-infrared and calculates as $(\text{NIR} + \text{Red})/(\text{NIR} + \text{Red})$)
    - indexes sometimes are similar to features extracted from time series data: ETCCDI climate change indices [@climate-indexes]: SU - number of summer days: annual count of days when TX > 25 degree
  - The index pipeline and software implementation we propose are applicable to all indexes mentioned above, while the main focus of this paper is on indexes that combine multiple variables/indexes since more skills are required to diagnose those indexes to understand how they behave.

### 2023-07-04 tidyindex.tex
outline for introducing done

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne, VIC, Australia
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

Many concepts relevant to decision making cannot be directly measured,
however, they are crucial for resource allocation, early prevention, and
other operational purpose. For example, fire authorities would be
interested to quantify fire risk since bushfires can have a huge impact
on monetary loss, health, and the local ecosystem. Climatologists would
be interested in monitoring the change in global climate since
variability in atmospheric and oceanic conditions has a direct impact on
global weather and climate. Usually this concept of interest is
associated with more than one variables and these variables need to be
integrated to make decisions on the subject matter. A common approach to
quantify concepts like these is to construct an index using these
relevant variables. This allows researchers to compare the quantity of
interest across entities (i.e.~countries, regions) and also cross time.

@@

In this article, an index is defined as a tool to quantify a concept of
interest that does not have a direct measure. The concept of interest
doesn't have a direct measure can because it is impractical to measure
at the population level. For example, it would be nearly impossible to
include all the available stocks in the market to characterise stock
market behavior, so indexes like Dow Jones Industrial Average, S\&P 500,
and Nasdaq Composite select a representative set of stocks to measure
the overall market behavior. Also belonging to this category are the
economic indexes like the Consumer Price Index, where price changes of a
basket of items are weighted to measure inflation. The lack of direct
measure could also because the concept itself is an unobservable human
construction, rather than a physical quantity that can be measured. Many
natural hazard and social concepts falls into this category. This
includes drought indexes constructed from meteorological, agricultural,
hydrological, and social-economic variables, e.g.~Standardised
Precipitation Index (SPI) (McKee et al. 1993) and Aggregated Drought
Index (ADI) (Keyantash and Dracup 2004) among others. Social development
indexes like Human Development Index (United Nations Development
Programme 2022) and Global Liveability Index (Economist Intelligence
Unit 2019) measure various aspects of the quality of human capital and
urban life.

\emph{still need to tweak the tone a bit: ``they are called index, they
are not the index we will talk about''}

Despite many quantity having the term \emph{index} in their name, they
cannot be technically classified as indexes according to the definition
given above. The reason for these quantities to lose their index
memberships is that they are variables can be accurately measured given
the instrument precision. This includes quantities like precipitation of
the driest month or percentage of days when maximum temperature is below
10th percentile. They are measures of precipitation and percentage of
days under specific conditions (dries month, maximum temperature below
10th percentile). They are variables, or indicators, that can be used to
construct indexes but are not indexes themselves. Similarly, a set of
remote sensing indexes are not indexes, since they are measures of
electromagnetic wave reflectance. This includes Normalized Difference
Vegetation Index (NDVI) (Tucker 1979), derived from the ratio of
difference over sum on two segments in the spectrum, also called band:
near-infrared (NIR) and red. So are the ``indexes'' derived from NDVI,
e.g.~Vegetation Condition Index (Kogan 1995). Notice that this does not
exclude all the construction derived from remotes sensor variables to be
valid indexes. For example, Vegetation Drought Response Index (Brown et
al. 2008) is a valid index since it integrates climate, satellite, and
biophysical variables to quantify vegetation stress.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-brown_vegetation_2008}{}}%
Brown, Jesslyn F., Brian D. Wardlow, Tsegaye Tadesse, Michael J. Hayes,
and Bradley C. Reed. 2008. {``The {Vegetation} {Drought} {Response}
{Index} ({VegDRI}): {A} {New} {Integrated} {Approach} for {Monitoring}
{Drought} {Stress} in {Vegetation}.''} \emph{GIScience \& Remote
Sensing} 45 (1): 16--46.
\url{https://doi.org/10.2747/1548-1603.45.1.16}.

\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

\leavevmode\vadjust pre{\hypertarget{ref-gli}{}}%
Economist Intelligence Unit. 2019. {``The Global Liveability Index
2019.''} The Economist.
\url{https://www.cbeinternational.ca/pdf/Liveability-Free-report-2019.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-keyantash_aggregate_2004}{}}%
Keyantash, John A., and John A. Dracup. 2004. {``An Aggregate Drought
Index: {Assessing} Drought Severity Based on Fluctuations in the
Hydrologic Cycle and Surface Water Storage.''} \emph{Water Resources
Research} 40 (9). \url{https://doi.org/10.1029/2003WR002610}.

\leavevmode\vadjust pre{\hypertarget{ref-kogan_application_1995}{}}%
Kogan, F. N. 1995. {``Application of Vegetation Index and Brightness
Temperature for Drought Detection.''} \emph{Advances in Space Research},
Natural {Hazards}: {Monitoring} and {Assessment} {Using} {Remote}
{Sensing} {Technique}, 15 (11): 91--100.
\url{https://doi.org/10.1016/0273-1177(95)00079-T}.

\leavevmode\vadjust pre{\hypertarget{ref-mckee1993relationship}{}}%
McKee, Thomas B, Nolan J Doesken, John Kleist, et al. 1993. {``The
Relationship of Drought Frequency and Duration to Time Scales.''} In
\emph{Proceedings of the 8th Conference on Applied Climatology},
17:179--83. 22. Boston, MA, USA.

@@

\leavevmode\vadjust pre{\hypertarget{ref-hdi}{}}%
---------. 2022. {``Human Development Report 2021-22.''} New York.
\url{http://report.hdr.undp.org}.

@@

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne, VIC, Australia
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

Many concepts relevant to decision making cannot be directly measured,
however, they are crucial for resource allocation, early prevention, and
other operational purpose.

\begin{itemize}
\tightlist
\item
  USDM and European Drought Observatory publishes drought indicator to
  {[}\ldots{]},
\item
  NOAA and BOM producing forecast on El Nino alert system to
  {[}\ldots{]}
\item
  OECD and UN producing indexes measuring socio-economic aspects of
  country for policy making.
\end{itemize}

There decision making often needs to consider multiple aspects and
indexes is a statistical tool to pack these information into a single
number, making it easier for policy makers and public communication.

@@

\begin{itemize}
\tightlist
\item
  Indexes can be composed from single variables or multiple variables
  (composite index)
\item
  to make the definition even more complex is that indexes themselves
  are used with other variables to compose indexes (many drought indexes
  using an index called SPI).
\item
  Some indexes may be more simple than you think:

  \begin{itemize}
  \tightlist
  \item
    a single transformation (NDVI (Tucker 1979) takes two spectrum: red
    and near-infrared and calculates as
    \((\text{NIR} + \text{Red})/(\text{NIR} + \text{Red})\))
  \item
    indexes sometimes are similar to features extracted from time series
    data: ETCCDI climate change indices (Zhang 2020): SU - number of
    summer days: annual count of days when TX \textgreater{} 25 degree
  \end{itemize}
\item
  The index pipeline and software implementation we propose are
  applicable to all indexes mentioned above, while the main focus of
  this paper is on indexes that combine multiple variables/indexes since
  more skills are required to diagnose those indexes to understand how
  they behave.
\end{itemize}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-climate-indexes}{}}%
Zhang, Xuebin. 2020. {``ETCCDI Climate Change Indices.''}
\url{http://etccdi.pacificclimate.org/index.shtml}.
